\textcolor[rgb]{0,0,1}{Data-Driven Model Predictive Control using Random Forests for Building Energy Optimization and Climate Control}

### 2018-01-22 paper/final/appendixRegTrees.tex
Sec 3 - DPC reviewed, and partially updated response to reviewers
prefinal edits
reviewed Intro, Sec 2,3,4,5, conclusion and appendix

\section{Regression Trees}
\label{A:RegTrees}
\textcolor[rgb]{0,0,1}{
In this appendix we explain how Regression Trees are built using an example adapted from~\cite{hastie2009elements}.
Tree-based methods partition the feature space into a set of rectangles (more formally, hyper-rectangles) and then fit a simple model in each one. They are conceptually simple yet powerful.
Let us consider a regression problem with continuous response $\Y = \{Y\}$ and and 2 predictors $\X = \{X_1,X_2\}$, each taking values in the unit interval. The top left plot of Figure~\ref{fig:friedman} shows a partition of the feature space by lines that are parallel to the coordinate axes. In each partition element, we can model $Y$ with a different constant.
\begin{figure}
  \centering
  \includegraphics[width=0.7\columnwidth]{figures/friedman1.eps}
  \vspace{-8pt}
  \caption{Top right: 2D feature space by recursive binary splitting. Top left: partition that cannot be obtained from recursive binary splitting. Bottom left: tree corresponding to the partition. Bottom right:  perspective plot of the prediction surface.}
  \label{fig:friedman}
   \vspace{-10pt}
\end{figure} 
However, there is a problem: although each partitioning line has a simple description like $X_1 = k$, some of the resulting regions are complicated to describe.
To simplify things, we can restrict ourselves to only consider recursive binary partitions, like the ones shown in the top right plot of Figure~\ref{fig:friedman}.
We first split the space into two regions, and model the response by the mean of $Y$ in each region. 
We choose the variable and split-point to achieve the best prediction for $Y$. 
Then one or both of these regions are split into two more regions, and this process is continued, until some stopping rule is applied. This is the "recursive partitioning" part of the  algorithm. 
For example, in the top right plot of Figure~\ref{fig:friedman}, we first split at $X_1 = t_1$. 
Then the region $X_1 \leq t_1$ is split at $X_2 = t_2$ and the region $X_1 > t_1$ is split at $X_1 = t_3$.
Finally, the region $X_1 > t_3$ is split at $X_2 = t_4$. 
The result of this process is a partition of the data-space into the five regions (or leaves) $R_1, R_2,\cdots, R_5$.
The corresponding regression tree model, $\T$, predicts $Y$ with a constant, $c_i$, in region $R_i$ i.e.,
\begin{equation}
\T = \mathit{f}_{\mathrm{tree}}(X_1,X_2) = \sum_{i=1}^{5} c_i I\left\lbrace(X_1,X_2) \in R_i\right\rbrace,
\label{eq:response}
\end{equation}
where $I\{X\in R\}$ is a function that is equal to $1$, if $X\in R$, and $0$, otherwise.
This same model can be represented by the binary tree shown in the bottom left of Figure~\ref{fig:friedman}. 
The full data-set sits at the top of the tree. 
Observations satisfying the condition at each node are assigned to the left branch, and the others to the right branch. The terminal nodes or leaves of the tree correspond to the regions $R_1,R_2,...,R_5$. 
The bottom right plot of Figure~\ref{fig:friedman}, shows the perspective plot of the regression surface obtained as a result of building a regression tree with the 5 constants $c_i,\ i=1,2,3,4,5$. }

\textcolor[rgb]{0,0,1}{
\paragraph{\textbf{Node splitting criteria}}
Now, a first question is: \emph{how to grow a regression tree?}
Suppose our dataset, $(\X,\Y)$, consisting of $p$ features, i.e. $\X = \{X_1,X_2,\ldots,X_p\}$, and one response variable, i.e. $\Y = \{Y\}$. Suppose we have $|(\X,\Y)| = n$ observations (samples): $(x_i,y_i)$, $i=1,2,\ldots,n$, with $x_i=(x_{i1},x_{i2},\ldots,x_{ip})$.
For regression trees we adopt the sum of squares as our splitting criteria, i.e. a variable at a node will be split if it minimizes the following sum of squares between the predicted response and the actual output variable:
\begin{equation}
\sum_i (y_i - \T(x_i))^2.
\end{equation}
The best response $c_i$ (from equation~\ref{eq:response} for the partition $R_i$), is just the average of output samples in the region $R_i$, i.e.
\begin{equation}
c_i = avg(y_i|x_i \in R_i).
\label{eq:average}
\end{equation}
Finding the best binary partition in terms of minimum sum of squares is generally computationally infeasible. 
A greedy algorithm is used instead. Starting with all of the data, consider a splitting variable $j$ and split point $s$, and define the following pair of left ($R_L$) and right ($R_R$) half-planes
\begin{equation}
\begin{aligned}
R_L(j,s) = \left\lbrace X|X_j \leq s \right\rbrace ,\\
R_R(j,s) = \left\lbrace X|X_j > s \right\rbrace
\end{aligned}
\end{equation}
The splitting variable $j$ and the split point $s$ is obtained by solving the following minimization:
\begin{equation}
\begin{aligned}
\underset{j,s}{\text{min}}\left[\underset{c_L}{\text{min}} \sum_{x_i\in R_L(j,s)} (y_i - c_L)^2 + \underset{c_R}{\text{min}} \sum_{x_i\in R_R(j,s)} (y_i - c_R)^2 \right]  
\end{aligned}
\label{eq:split}
\end{equation}
where, for any choice of $j$ and $s$, the inner minimization in equation~\ref{eq:split} is solved using
\begin{equation}
\begin{aligned}
c_L = \text{avg}(y_i|x_i \in R_L(j,s)),\\
c_R = \text{avg}(y_i|x_i \in R_R(j,s)).
\end{aligned}
\end{equation}
For each splitting variable $X_j$, the determination of the split point $s$ can be done very quickly and hence by scanning through all of the inputs ($X_i$'s), the determination of the best pair $(j, s)$ is feasible.
Having found the best split, we partition the data into the two resulting regions and repeat the splitting process on each of the two regions. Then this process is repeated on all of the resulting regions.}

@@

\textcolor[rgb]{0,0,1}{
\paragraph{\textbf{Stopping criteria and pruning}}
At this point, the second question is: \emph{How large should we grow the tree?}
Every recursive algorithm needs to know when it's done, i.e. it requires a stopping criteria. 
For regression trees this means when to stop splitting the nodes. 
A very large tree might over fit the data, while a small tree might not capture the important structure.
Tree size is a tuning parameter governing the model’s complexity, and the optimal tree size should be adaptively chosen from the data. 
One approach is to split tree nodes only if the decrease in sum-of-squares due to the split exceeds some threshold. 
However, this strategy is myopic, since a seemingly worthless split might lead to a very good split below it.
A preferred, strategy is to grow a large tree, stopping the splitting process only when some minimum number of data points at a node (\texttt{MinLeaf}) is reached. Then this large tree is pruned using cost-complexity pruning methods.}

>>>>

\section{Regression Trees}
\label{A:RegTrees}
\textcolor[rgb]{0,0,1}{
In this appendix we explain how Regression Trees are built using an example adapted from~\cite{hastie2009elements}.
Tree-based methods partition the feature space into a set of rectangles (more formally, hyper-rectangles) and then fit a simple model in each one. They are conceptually simple yet powerful.
Let us consider a regression problem with continuous response $\Y = \{Y\}$ and 2 predictors $\X = \{X_1,X_2\}$, each taking values in the unit interval. The top left plot of Figure~\ref{fig:friedman} shows a partition of the feature space by lines that are parallel to the coordinate axes. In each partition element, we can model $Y$ with a different constant.
\begin{figure}
  \centering
  \includegraphics[width=0.7\columnwidth]{figures/friedman.eps}
  \vspace{-8pt}
  \caption{Top right: 2D feature space by recursive binary splitting. Top left: partition that cannot be obtained from recursive binary splitting. Bottom left: tree corresponding to the partition. Bottom right:  perspective plot of the prediction surface.}
  \label{fig:friedman}
   \vspace{-10pt}
\end{figure} 
However, there is a problem: although each partitioning line has a simple description like $X_1 = k$, some of the resulting regions are complicated to describe.
To simplify things, we can restrict ourselves to only consider recursive binary partitions, like the ones shown in the top right plot of Figure~\ref{fig:friedman}.
We first split the space into two regions, and model the response by the mean of $Y$ in each region. 
We choose the variable and split-point to achieve the best prediction for $Y$. 
Then one or both of these regions are split into two more regions, and this process is continued, until some stopping rule is applied. This is the "recursive partitioning" part of the  algorithm. 
For example, in the top right plot of Figure~\ref{fig:friedman}, we first split at $X_1 = t_1$. 
Then the region $X_1 \leq t_1$ is split at $X_2 = t_2$ and the region $X_1 > t_1$ is split at $X_1 = t_3$.
Finally, the region $X_1 > t_3$ is split at $X_2 = t_4$. 
The result of this process is a partition of the data-space into the five regions (or leaves) $R_1, R_2,\cdots, R_5$.
The corresponding regression tree model, $\T$, predicts $Y$ with a constant, $c_i$, in region $R_i$ i.e.,
\begin{equation}
\mathit{f}_{\T}(X_1,X_2) = \sum_{i=1}^{5} c_i I\left\lbrace(X_1,X_2) \in R_i\right\rbrace,
\label{eq:response}
\end{equation}
where $I\{X\in R\}$ is a function that is equal to $1$, if $X\in R$, and $0$, otherwise.
This same model can be represented by the binary tree shown in the bottom left of Figure~\ref{fig:friedman}. 
The full data-set sits at the top of the tree. 
Observations satisfying the condition at each node are assigned to the left branch, and the others to the right branch. The terminal nodes or leaves of the tree correspond to the regions $R_1,R_2,...,R_5$. 
The bottom right plot of Figure~\ref{fig:friedman}, shows the perspective plot of the regression surface obtained as a result of building a regression tree with the 5 constants $c_i,\ i=1,2,3,4,5$. }

\textcolor[rgb]{0,0,1}{
\paragraph{\textbf{Node splitting criteria}}
Now, a first question is: \emph{how to grow a regression tree?}
Suppose our dataset, $(\X,\Y)$, consisting of $p$ features, i.e. $\X = \{X_1,X_2,\ldots,X_p\}$, and one response variable, i.e. $\Y = \{Y\}$. Suppose we have $|(\X,\Y)| = n$ observations (samples): $(x_i,y_i)$, $i=1,2,\ldots,n$, with $x_i=(x_{i1},x_{i2},\ldots,x_{ip})$.
For regression trees we adopt the sum of squares as our splitting criteria, i.e. a variable at a node will be split if it minimizes the following sum of squares between the predicted response and the actual output variable:
\begin{equation}
\sum_i (y_i - \mathit{f}_{\T}(x_i))^2.
\end{equation}
The best response $c_i$ (from equation~\ref{eq:response} for the partition $R_i$), is just the average of output samples in the region $R_i$, i.e.
\begin{equation}
c_i = avg(y_i|x_i \in R_i).
\label{eq:average}
\end{equation}
Finding the best binary partition in terms of minimum sum of squares is generally computationally infeasible. 
A greedy algorithm is used instead. Starting with all of the data, consider a splitting variable $j$ and split point $s$, and define the following pair of left ($R_L$) and right ($R_R$) half-planes
\begin{equation}
\begin{aligned}
R_L(j,s) = \left\lbrace X|X_j \leq s \right\rbrace ,\\
R_R(j,s) = \left\lbrace X|X_j > s \right\rbrace
\end{aligned}
\end{equation}
The splitting variable $j$ and the split point $s$ is obtained by solving the following minimization:
\begin{equation}
\begin{aligned}
\underset{j,s}{\text{min}}\left[\underset{c_L}{\text{min}} \sum_{x_i\in R_L(j,s)} (y_i - c_L)^2 + \underset{c_R}{\text{min}} \sum_{x_i\in R_R(j,s)} (y_i - c_R)^2 \right]  
\end{aligned}
\label{eq:split}
\end{equation}
where, for any choice of $j$ and $s$, the inner minimization in equation~\ref{eq:split} is solved using
\begin{equation}
\begin{aligned}
c_L = \text{avg}(y_i|x_i \in R_L(j,s)),\\
c_R = \text{avg}(y_i|x_i \in R_R(j,s)).
\end{aligned}
\end{equation}
For each splitting variable $X_j$, the determination of the split point $s$ can be done very quickly and hence by scanning through all of the inputs ($X_i$'s), the determination of the best pair $(j, s)$ is feasible.
Having found the best split, we partition the data into the two resulting regions and repeat the splitting process on each of the two regions. Then this process is repeated on all of the resulting regions.}

@@

\textcolor[rgb]{0,0,1}{
\paragraph{\textbf{Stopping criteria and pruning}}
At this point, the second question is: \emph{How large should we grow the tree?}
Every recursive algorithm needs to know when it's done, i.e. it requires a stopping criteria. 
For regression trees this means when to stop splitting the nodes. 
A very large tree might over fit the data, while a small tree might not capture the important structure.
Tree size is a tuning parameter governing the model’s complexity, and the optimal tree size should be adaptively chosen from the data. 
One approach is to split tree nodes only if the decrease in sum-of-squares due to the split exceeds some threshold. 
However, this strategy is myopic, since a seemingly worthless split might lead to a very good split below it.
A preferred strategy is to grow a large tree, stopping the splitting process only when some minimum number of data points at a node (\texttt{MinLeaf}) is reached. Then this large tree is pruned using cost-complexity pruning methods.}

### 2018-01-22 paper/final/casestudyEPlus.tex
reviewed Intro, Sec 2,3,4,5, conclusion and appendix

\section{APPLICATION: DEMAND RESPONSE}
\label{S:casestudy}

@@

Thus, the problem of energy management during a DR event makes an ideal case for DPC. In the following sections, we apply DPC-En to a large scale EnergyPlus model to show how effectively DPC can provide a desired power curtailment as well as a desired thermal comfort. DPC builds predictive models of a building based on historical weather, schedule, set-points and electricity consumption data, while also learning from the actions of the building operator. These models are then used for synthesizing recommendations about the control actions that the operator needs to take, during a DR event, to obtain a given load curtailment while providing guarantees on occupant comfort and operations.

@@

In the following simulations, we consider a long DR event from 7am - 2pm when the end-users are expected to follow/track the reference power signal sent by the utility. This is indeed common in Demand Tracking Control. 
During offline training, we sample data every 15 min to learn 2 kinds of forests. (1) Power forests are built using output as the total building power consumption, and (2) Temperature forests with output as temperature of one of the 22 zones. 
The training data set contains the following types of features. (1) The \textit{weather data} which includes measurements of the outside air temperature and relative humidity. Since we are interested in predicting the power consumption or the zone temperature for a finite horizon, we include the weather forecast of the complete horizon in the training features. (2) The \textit{schedule data} includes the proxy variables which correlate with repeated patterns of electricity consumption e.g., due to occupancy or equipment schedules. Day of Week is a categorical predictor which takes values from 1-7 depending on the day of the week. This variable can capture any power consumption patterns which occur on specific days of the week. Likewise, Time of Day is quite an important predictor of power consumption as it can adequately capture daily patterns of occupancy, lighting and appliance use without directly measuring any one of them. Besides using proxy schedule predictors, actual building equipment schedules can also be used as training data for building the trees. (3) The \textit{building data} include (i) cooling set points for  the guest rooms, kitchen and corridors, (ii) supply air temperature, and (iii) chilled water temperature.
For the following simulations, we use five control variables (i) cooling set point for corridors $\mathsf{ClgSP}$, (ii) cooling set point for guest rooms $\mathsf{GuestSP}$, (iii) cooling set point for kitchen $\mathsf{KitchenSP}$, (iv) chilled water supply temperature $\mathsf{ChwSP}$, and (v) supply air temperature  $\mathsf{SupplyAirSP}$, so $u := [\mathsf{ClgSP},\mathsf{GuestClgSP},\mathsf{KitchenClgSP},\mathsf{SupplyAirSP}, \mathsf{ChwSP}]^\top$. 
The power forest $\mathcal{F}_p$ is built using the total building power consumption $\tP$. Its features in $\X^d$ include the weather variables, their lag terms and their forecast over the horizon, the schedule variables, and finally the lag terms of the power consumption.
The temperature forest $\mathcal{R}_t$ is built with zone temperature $\tT$ as the output. Except for the autoregressive terms corresponding to the same zone temperature, all other features are same in $\X^d$.
Figure~\ref{F:sepvars} shows the prediction accuracy for the power forest, and also explains the two level training approach introduced in Section~\ref{SS:sepvar}. During S1, the forests are trained using only disturbances as the features. Then in S2, the local effects of the control variables are accounted for by the linear models in the leaves. We observe how the accuracy is drastically improved after including the linear models in the predictions.

@@

\begin{figure}[h!]
	\begin{center}
	\subfigure[Optimal inputs calculated by DPC-En. At first, the inputs are changed rapidly because of a significant difference between the desired and the actual power consumption. Then gradual adjustments are made to follow the desired reference.]{
		\label{F:eplus_inputs}
		\centering
		\includegraphics[width=29pc]{figures/eplus_control.eps}
	}
	\subfigure[Power tracking by DPC-En at 1.1 MW. In blue is the response when the optimal input is applied to the power Random Forest and in red is the response when the same input is applied to the plant (EnergyPlus building). The difference in closed-loop simulation and prediction is due to model mismatch.]{
		\label{F:eplus_power}
		\centering
		\includegraphics[width=29pc]{figures/eplus_closedloop.eps}
	}
\end{center}
	\caption{Power management using DPC. The controller is active between 7am - 2pm. This region is marked in dashed red lines.}
	\captionsetup{justification=centering}
	\label{F:eplus_track}
\end{figure}

@@

The results are shown in Figure~\ref{F:eplus_track}. 
The DPC controller is active between 7am - 2pm. Before 7am and after 2pm, the building is using a predefined rule-based control strategy.
The optimal control inputs from DPC-En are shown in Figure~\ref{F:eplus_inputs}. It is observed that, with the optimal inputs generated by DPC, we can track the reference power consumption signal closely. In fact, the average tracking error between 7am - 2pm is 3\%.
\textcolor[rgb]{0,0,1}{The results from the closed-loop simulation are shown in Figure~\ref{F:eplus_power}.
Marked in blue is the response when the optimal input is applied to the power Random Forest and in red is the response when the same input is applied to the plant (EnergyPlus building).
The difference between the predicted power consumption and that in the closed-loop simulation is due to model mismatch between the EnergyPlus model and the power forest used in the optimization \eqref{E:ptrack}.} 
Due to this inaccuracy, the actual power consumption is on an average 7 kW higher.

>>>>

\section{APPLICATION TO DEMAND RESPONSE}
\label{S:casestudy}

@@

Thus, the problem of energy management during a DR event makes an ideal case for DPC. In the following sections, we apply DPC-En to a large scale EnergyPlus model to show how effectively DPC can provide a desired power curtailment as well as a desired thermal comfort. DPC builds predictive models of a building based on historical weather, schedule, set-points and electricity consumption data, while also learning from the actions of the building operator. These models are then used for synthesising recommendations about the control actions that the operator needs to take, during a DR event, to obtain a given load curtailment while providing guarantees on occupant comfort and operations.

@@

In the following simulations, we consider a long DR event from 7am - 2pm when the end-users are expected to follow/track the reference power signal sent by the utility. This is indeed common in Demand Tracking Control. 
During offline training, we sample data every 15 min to learn 2 kinds of forests. (1) Power forests are built using output as the total building power consumption, and (2) Temperature forests with output as temperature of one of the 22 zones. 
The training data set contains the following types of features. (1) The \textit{weather data} which includes measurements of the outside air temperature and relative humidity. Since we are interested in predicting the power consumption or the zone temperature for a finite horizon, we include the weather forecast of the complete horizon in the training features. (2) The \textit{schedule data} includes the proxy variables which correlate with repeated patterns of electricity consumption e.g., due to occupancy or equipment schedules. Day of Week is a categorical predictor which takes values from 1-7 depending on the day of the week. This variable can capture any power consumption patterns which occur on specific days of the week. Likewise, Time of Day is quite an important predictor of power consumption as it can adequately capture daily patterns of occupancy, lighting and appliance use without directly measuring any one of them. Besides using proxy schedule predictors, actual building equipment schedules can also be used as training data for building the trees. (3) The \textit{building data} include (i) cooling set points for  the guest rooms, kitchen and corridors, (ii) supply air temperature, and (iii) chilled water temperature.
For the following simulations, we use five control variables (i) cooling set point for corridors $\mathsf{ClgSP}$, (ii) cooling set point for guest rooms $\mathsf{GuestSP}$, (iii) cooling set point for kitchen $\mathsf{KitchenSP}$, (iv) chilled water supply temperature $\mathsf{ChwSP}$, and (v) supply air temperature  $\mathsf{SupplyAirSP}$, so $u := [\mathsf{ClgSP},\mathsf{GuestClgSP},\mathsf{KitchenClgSP},\mathsf{SupplyAirSP}, \mathsf{ChwSP}]^\top$. 
The power forest $\mathcal{F}_p$ is built using the total building power consumption $\tP$. Its features in $\X^d$ include the weather variables, their lag terms and their forecast over the horizon, the schedule variables, and finally the lag terms of the power consumption.
The temperature forest $\mathcal{F}_t$ is built with zone temperature $\tT$ as the output. Except for the autoregressive terms corresponding to the same zone temperature, all other features are same in $\X^d$.
Figure~\ref{F:sepvars} shows the prediction accuracy for the power forest, and also explains the two-steps training approach introduced in Section~\ref{S:dpc}. During S1, the forests are trained using only disturbances as the features. Then in S2, the local effects of the control variables are accounted for by the linear models in the leaves. We observe how the accuracy is drastically improved after including the linear models in the predictions.

@@

\begin{figure}[h!]
	\begin{center}
	\subfigure[Optimal inputs calculated by DPC-En. At first, the inputs are changed rapidly because of a significant difference between the desired and the actual power consumption. Then gradual adjustments are made to follow the desired reference.]{
		\label{F:eplus_inputs}
		\centering
		\includegraphics[width=29pc]{figures/eplus_control.eps}
	}
	\subfigure[Power tracking by DPC-En at 1.1 MW. In blue is the response when the optimal input is applied to the power predictive model of the Random Forest and in red is the response when the same input is applied to the power predictive model of EnergyPlus). The difference in closed-loop simulation and prediction is due to model mismatch.]{
		\label{F:eplus_power}
		\centering
		\includegraphics[width=29pc]{figures/eplus_closedloop.eps}
	}
\end{center}
	\caption{Power management using DPC. The controller is active between 7am - 2pm. This region is marked in dashed red lines.}
	\captionsetup{justification=centering}
	\label{F:eplus_track}
\end{figure}

@@

The results are shown in Figure~\ref{F:eplus_track}. 
The DPC controller is active between 7am - 2pm. Before 7am and after 2pm, the building is using a predefined rule-based control strategy.
The optimal control inputs from DPC-En are shown in Figure~\ref{F:eplus_inputs}. It is observed that, with the optimal inputs generated by DPC, we can track the reference power consumption signal closely. In fact, the average tracking error between 7am - 2pm is 3\%.
\textcolor[rgb]{0,0,1}{The results from the closed-loop simulation are shown in Figure~\ref{F:eplus_power}.
Marked in blue is the response when the optimal input is applied to the power predictive model of the Random Forest \eqref{E:ptrack} and in red is the response when the same input is applied to the power predictive model of EnergyPlus. Since the optimal input is computed using the power predictive model of the Random Forest the blue trajectory perfectly follows the tracking signal. The red trajectory, as expected, is characterised by a (small) tracking error because of the model mismatch between the predictive model of the Random Forest (used to compute the input) and that of EnergyPlus (used to simulate the closed-loop system).} 
Due to this inaccuracy, the actual power consumption is on an average 7 kW higher.

### 2018-01-22 paper/final/conclusion.tex
reviewed Intro, Sec 2,3,4,5, conclusion and appendix

To overcome the difficulties associated with the model identification in Model Predictive Control (MPC), we introduce a novel idea for predictive control using data: Data-driven model Predictive Control (DPC). \textcolor[rgb]{0,0,1}{Data-driven control is based on non-physical (black-box) models, therefore they can not be integrated with most of the classical control approaches.} The goal is to create data-driven models that are suitable for receiding horizon control. \textcolor[rgb]{0,0,1}{To this aim, we present two algorithms, based on trees and random forests, to create control-oriented models for DPC. We then apply DPC to three different case studies to demonstrate its strength.}
\begin{enumerate}
\item \emph{\textbf{Comparison with MPC.}} We compare the performance of our DPC to MPC on a multivariable bilinear building model. We establish that DPC with random forests shows a remarkable similarity to MPC in the optimal control strategies explaining $70\%$ variance. On the other hand, DPC with regression trees suffers from practical limitations due to model overfitting.
\item \emph{\textbf{Demand Response application.}} We further apply DPC with random forests to a large scale 6 story EnergyPlus model with 22 zones for which the traditional model-based control is largely unsuitable due to complex dynamics and the cost of model identification. We show that DPC, relying only on the sensor data, can provide significant energy savings while maintaining thermal comfort. Our results demonstrate that even for such complex system, DPC tracks a reference signal with a mean error of $3\%$.
\item \textcolor[rgb]{0,0,1}{\emph{\textbf{Optimal heating system scheduling of a real house.}} We finally apply DPC using experimental data from a real house and compare it with the classical bang-bang controller widely used for temperature control. We show that DPC provides energy savings while guaranteeing thermal comfort for the occupants. With respect to a bang-bang controller, we obtain energy savings that go from $25.4\%$, when we force the temperature to be strictly in a comfort range, to $49.2\%$, if we allow small violations in the thermal comfort level. Furthermore, we demonstrate the robustness of DPC considering uncertainties in the weather forecast.}
\end{enumerate}

DPC has applications which go beyond buildings and energy systems, to industrial process control, and controlling large critical infrastructures like water networks, district heating \& cooling. DPC is immensely valuable in situations where first principles based modeling cost is extremely high.

>>>>

To overcome the difficulties associated with the model identification in Model Predictive Control (MPC), we introduce a novel idea for predictive control using data: Data-driven model Predictive Control (DPC). \textcolor[rgb]{0,0,1}{Data-driven control is based on non-physical (black-box) models, therefore they can not be integrated with most of the classical control approaches.} The goal is to create data-driven models that are suitable for receiding horizon control. \textcolor[rgb]{0,0,1}{To this aim, we present two algorithms, based on trees and random forests, to create control-oriented models for DPC. We then apply DPC to three different case studies to demonstrate its strength.}
\begin{enumerate}
\item \emph{\textbf{Comparison with MPC.}} We compare the performance of our DPC to MPC on a multivariable bilinear building model. We establish that DPC with random forests shows a remarkable similarity to MPC in the optimal control strategies explaining $70\%$ variance. On the other hand, DPC with regression trees suffers from practical limitations due to model overfitting.
\item \emph{\textbf{Application to Demand Response.}} We further apply DPC with random forests to a large scale 6 story EnergyPlus model with 22 zones for which the traditional model-based control is largely unsuitable due to complex dynamics and the cost of model identification. We show that DPC, relying only on the sensor data, can provide significant energy savings while maintaining thermal comfort. Our results demonstrate that even for such complex system, DPC tracks a reference signal with a mean error of $3\%$.
\item \textcolor[rgb]{0,0,1}{\emph{\textbf{Application to optimal heating system scheduling.}} We demonstrate robustness of our method to uncertainties due to real data acquisition and weather forecast inaccuracies by implementing and testing DPC on historical data from an off-grid house located in L'Aquila, Italy. We derive a predictive model on such real data and design the optimal ON/OFF scheduling for the heating system in order to save energy while guaranteeing thermal comfort for the occupants. We compare the total amount of energy saved with respect to the classical bang-bang controller (widely used in houses for temperature control) using an EnergyPlus model built specifically for the house. We show that we can perform an energy saving that ranges from $25.4\%$ (if we guarantee thermal comfort i.e.~strictly respect the desired temperature range in the rooms) to $49.2\%$ (if we allow small violation in the desired temperature range). Finally, we test the robustness of our method to uncertainties in the real data acquisition and weather forecast.}
\end{enumerate}

DPC has applications which go beyond buildings and energy systems, to industrial process control, and controlling large critical infrastructures like water networks, district heating \& cooling. In general, DPC is immensely valuable in situations where first principles based modeling cost is extremely high.
